|

### 2021-07-27 03-Chp-Systematic-Review.Rmd
End of day:

### Search strategy

@@

### Study selection

@@

@@

<!----------------------------------------------------------------------------->
#### Non-randomised studies of interventions/exposures

@@

As part of this review, I piloted the RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool. <!-- TODO CITATION NEEDED --> Results were visualise

### Studification

### Following up with authors {#contacting-authors}

@@

@@

@@

<!-- TODO Need equation for random effects too -->

@@

### Visualisation of results

@@

@@

@@

Patient and public memebers were intended to be involved at this poit

@@

@@

@@

### Converting risk ratios to odds ratios

@@

Straight-forward random effects meta-analysis of

@@

@@

A common criticism of preprinted articles is that they have not formally been peer-reviewed. __[CITATION NEEDED]__ However, I believe this criticism is less warranted in the context of inclusion of preprints in a systematic review, given a formal assessment of risk of bias, which provides a structured approach to assessing the internal validity of a study.

@@

While [__some__] preprint have subsequently been published, this does not cause any major issues.

@@

@@

@@

### Original contribution to research

@@

### Comparison with other reviews

@@

#### Inclusion of Mendelian randomisation studies

@@

>>>>

### Search strategy and and study selection

@@

#### Study selection

@@

#### Studification

#### Following up with authors {#contacting-authors}

#### Converting between different

WHile the analysis of change scores is feasible, as they were explicityl

@@

#### Non-randomised studies of interventions/exposures

@@

As part of this review, I piloted the RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool. <!-- TODO CITATION NEEDED --> Results were visualised

@@

Studies were not combined across different study designs (i.e. RCTs were not combined in a meta-analysis with results from observational studies).

@@

$u$

\begin{equation}
  \theta_i = \mu + u_i
  (\#eq:meta-analysis-random)
\end{equation}

@@

<!-- TODO Need equation for random effects too. See Sean's thesis here. -->

This analytical approach was used for cases where the exposure was dichotomous (e.g. statin use or not, hypercholesterolemia or not).

@@

#### Visualisation of results

@@

Also, the use of evidence maps is becoming increasingly common to explore the distibution of research cohorts included in a

As such, each individual cohort contibuting to the evidence base was quantified and visualised on a world map.

@@

<!-- TODO List here that one of the key reasons for building the tool was to . Reference back to  -->

@@

Patient and public members were intended to be involved at this poit

@@

Similarly, there was initially no intention to try and examine the risk of bias due to missing evidence.

@@

<!-- IDEA Could include a second evidence synthesis map here to summarise Table 1, highlighting the lack of vascular dementia outcomes and the massive amount of observational data -->

@@

@@

Straight-forward random effects meta-analysis of binary exposures

@@

Important here to examine effect of risk of bias (due to immortal time in observational studies).

Also important to examine the effect of:

* age at baseline
* sex (percentage)
* education
* baseline cognitive scores (how to group)

@@

A common criticism of preprinted articles is that they have not formally been peer-reviewed. __[CITATION NEEDED]__ <!-- TODO See Chapter 1 for reference for this quote. -->
 However, I believe this criticism is less warranted in the context of inclusion of preprints in a systematic review, given a formal assessment of risk of bias, which provides a structured approach to assessing the internal validity of a study.

@@

While `r hold()` preprint have subsequently been published, this does not cause any major issues.

### Risk of bias due to missing evidence

@@

### Summary of findings

@@

In retrospective, an alternative approach to the one taken could have been employed, performing an review of reviews, also known as an umbrella review. In these studies, the unit of analysis are systematic reviews rather than the

However, a comparative analysis between indicates

@@

@@

#### Limitations

@@

### Strengths and limitations

@@

In this chapter I have

This work built on the tool introduced in the preceeding chapter, and the results from the systematic review are used in further chapters.

In Chapter \@ref(cprd-analysis-heading), 
A summary of the evidence guided the

In Chapter \@ref(ipd-heading),

<!-- IDEA Link here to IPD, talking about how the meta-regression on key variables suggested the use of -->

A

### 2021-07-27 manuscripts/cprd/manuscript.Rmd
End of day:

__Methods:__ A cohort study design using data from the CPRD, routinely collected between January 1995 and March 2016, was performed. Cox proportional hazard models, allowing for a time-varying treatment indicator, were used to estimate the association between seven lipid-regulating drug classes and five dementia outcomes (all-cause, vascular and other dementia, and probable and possible AD).

__Results:__ We analyzed `r totaltext` with a total follow-up of `r comma(total_followup)` participant-years (median: `r fu_text`). We found little evidence that lipid-regulating agents were associated with risk of AD (probable AD `r probad_text`; possible AD `r possad_text`), but there was evidence of an increased risk of all-cause (`r anydem_text`), vascular (`r vasdem_text`) and other dementia (`r othdem_text`). Evidence from a number of control outcomes (ischeamic heart disease `r ihd_text`; backpain `r backpain_text` and diabetes `r dm_type2_text`) indicated the presence of substantial residual confounding related to vascular factors.

@@

* Little evidence that lipid-regulating agents were associated with Alzheimer’s disease, but there was some evidence for a harmful effect on all-cause, vascular and other dementia. In all cases, the estimated associations were driven by the statin subgroup, which comprised most participants in our cohort.

@@

Dementia is a major neurocognitive disorder, the most common types of which are Alzheimer’s disease (AD) and vascular dementia (VaD).[@prince2016] Despite an increasing number of cases globally and decades of research, there remains much unknown about the pathogenesis and progression of the disease, and, at present, no effective treatment exists to arrest or reverse the cognitive decline associated with the condition.[@jobke2018] Drug repurposing, the identification of new applications for previously approved drugs, may provide an efficient mechanism to discover effective preventative and therapeutic treatments for dementia.[@ashburn2004; @pushpakom2019]

@@

We therefore aim to examine the association between several major classes of LRA and all-cause dementia, Alzheimer's disease, vascular dementia and other dementia, in the Clinical Practice Research Datalink (CPRD), a large, population-based electronic health record (EHR) database.[@herrett2015]

@@

We performed a cohort study using data from the CPRD. Our initial sample included all participants registered at a participating practice between 1 January 1995 and 29 February 2016 who had a flag for "research quality" data. Records pre-dating the 1995 cutoff were excluded from the analysis as data quality and reliability is thought to be higher after this date.[@wolf2019] All events of interest were identified using predetermined code lists, which are available for inspection (see [Data/code availability](#data-code-avail)).

@@

Participants were included in our study cohort if their record contained any of the following index events: a Read code for a diagnosis of hypercholesterolemia or related condition; a Read code for prescription of a LRA (such as statins); a total cholesterol test result of >4mmol/L; or an low density lipoprotein cholesterol (LDL-c) test result of >2mmol/L.

These index events allowed us to define a population of participants who were either at risk of hypercholesterolemia, as indicated by the elevated total or LDL-c test results, or had already been diagnosed with it, as indicated by a diagnostic code or related prescription. This approach, conditioning entry into the study on being either “at-risk” or already diagnosed with hypercholesterolemia, was employed in an attempt to reduce confounding by indication that we would expect to observe in a general population cohort.

The index date for a participant was defined as the date where the first relevant code or test result (as detailed above) was recorded on their clinical record. Participants were followed up until the earliest of: an outcome of interest; death; end of follow-up (29 February 2016); or last registration date with their GP practice. Participants were removed from our sample if they were less than 40 years of age, had less than 12 months of "research quality" data, were simultaneously prescribed more than one LRA (due to the difficulty of assigning these patients to a single exposure group), or were diagnosed with dementia before or on the date of the index event.

@@

A participant’s drug class was assigned based on their first recorded prescription, and any drug switching was ignored in an effort to mimic an intention-to-treat approach. We did however tabulate how often the initial drug class was stopped (defined as last prescription of the primary class being followed by at least six months of observation), added to (defined as a second drug class being prescribed before the last prescription of the initial class), or switched (defined as a second drug class being prescribed after the last prescription of the initial class).

@@

All analyses were performed in STATA 15. Cox proportional hazard models were used to estimate the hazard ratio and corresponding 95% confidence intervals, allowing for the potential for clustering by practice. Participant's age was used as the time axis for all models.[@lamarca1998; @gail2009; @pencina2007] To address the potential for immortal time bias, we employed a time-varying indicator of treatment status to correctly allocate time-at-risk to the exposed and unexposed groups.[@levesque2010] To observe the effect of adjusting for additional covariates, we compared models adjusted for age only and age and sex to the fully adjusted model. Additional analyses stratified by outcome and drug class were also performed.

@@

## Additional analyses

We performed a number of sensitivity analysis. As statins are contraindicated in pregnancy,[@nationalinstituteforhealthandcareexcellence]  we ran the models described above but excluding participants below the age of 55. We also further stratified the statin exposure group into lipophilic (Atorvastatin, Lovastatin, Simvastatin, Cerivastatin) and hydrophilic (Pravastatin, Rosuvastatin, Fluvastatin) statins. Finally, we included three control outcomes with known associations with statin use.[@collins2016a; @herrett2021] Using the fully adjusted model, we investigated the association between LRA and back pain (negative control), ischemic heart disease (positive protective control), and Type 2 diabetes (positive harmful control).

@@

A total of `r format(attrition[14,1], big.mark = ",")` participants met the inclusion criteria for our cohort (See Supplementary Figure 2 for the attrition flowchart), with a total follow-up of `r comma(total_followup)` patient years at risk. The majority of participants were included in the cohort on the basis of an elevated test result (`r index_event_text`). The median age at index was `r age_text` and participants were followed up for a median of `r fu_text`. During follow-up, an all-cause dementia diagnosis was recorded for `r format(p1[1,2], big.mark = ",")` patients  (`r format(p1[2,2], big.mark = ",")` probable AD, `r format(p1[3,2], big.mark = ",")` possible AD, `r format(p1[4,2], big.mark = ",")` vascular dementia, `r format(p1[5,2], big.mark = ",")` other dementia). The distribution of baseline characteristics across the drug classes can be seen in Table \@ref(tab:cprdCharacteristics-table).

A substantial majority (`r percentage.statins`) of participants prescribed a lipid-regulating agent were prescribed a statin. We excluded the "Ezetimibe and statins" and "Nicotinic acid groups" classes from subsequent analysis based on the extremely small number of participants in these groups (Table \@ref(tab:cprdCharacteristics-table)). The stopping, addition and switching of drug classes was common across all exposure groups (Supplementary Table 3).

@@

As shown in Figure \@ref(fig:cprdPrimary), our results show little evidence of an effect of lipid-regulating agents on probable (`r probad_text`) and possible (`r possad_text`) Alzheimer's disease when compared with no treatment, with the exception of fibrates on probable Alzheimer's disease (`r probad_fib_text`).

@@

In contrast to the findings for Alzheimer's disease outcomes, lipid-regulating agents were associated with an increased risk of a subsequent diagnosis of vascular dementia (`r vasdem_text`) or other dementia (`r othdem_text`). Again this effect was driven mainly by the statin subgroup, but there was some evidence that ezetimibe was associated with an increased risk of vascular (`r vasdem_eze_text`) and other (`r othdem_eze_text`) dementia.

@@

Adjustment for additional covariates beyond age and sex had a limited impact (Supplementary Figure 3), with the exception of the Probable AD outcome, where the full adjustment attenuated to the null the protective effect observed when adjusting only for age and sex.

@@

For our control outcomes (Supplementary Figure 6), there was some evidence that treatment with a lipid-regulating agent was associated with an increased risk of back pain (`r backpain_text`). However, LRA prescription was also associated with an increased risk of ischemic heart disease (`r ihd_text`) and Type 2 diabetes (`r dm_type2_text`).

@@

There was little evidence that lipid-regulating agents had any observed effect on probable and possible Alzheimer's when compared with no treatment, but some evidence they were associated with an increased risk of an all-cause dementia, vascular dementia and other dementia diagnosis. The effect observed in each case was driven by the statin subgroup, which included a substantial majority of participants. For the other drug classes, there was limited evidence of an association with any outcome, with two exceptions. Ezetimibe was associated with increased risk of vascular and other dementia, while fibrates were associated with increased risk of all-cause dementia and probable Alzheimer's disease.

@@

These sources of evidence conflict with the findings of our analysis, where statin use was associated with an increased risk of all-cause dementia (`r anydem_text`). However, some of the included studies in the meta-analysis specifically exclude vascular dementia from the definition of all-cause dementia,[@chao2015] which may lead to an artificial protective effect of statins on all-cause dementia.

@@

Our results are broadly in line with the findings of two distinct approaches examining the effect of statin treatment on subsequent Alzheimer's disease. No randomized trials of statins for the prevention of Alzheimer's disease have been reported, but a recent meta-analysis of 20 observational studies found statins were associated with a reduced risk of Alzheimer's disease (RR 0.69, 95% CI 0.60–0.80) with stronger evidence than observed in our analysis.[@poly2020] In addition, a recent Mendelian randomization study examining the effect of genetic inhibition of HMGCR on Alzheimer’s disease found a small reduction in risk of Alzheimer's disease, comparable in magnitude to our findings (OR: 0.91, 95%CI: 0.63-1.31).[@williams2020]

An additional analysis found no difference in effect between lipophilic and hydrophilic statins for the prevention of Alzheimer's disease, consistent with a recent meta-analysis.[@chu2018]

@@

Much less literature is available on the association between lipid-regulating agents and vascular dementia or other dementia. A recent review found four observational studies examining the association of statins and vascular dementia found limited evidence for an effect (RR:0.93, 95% CI 0.74–1.16).[@poly2020] This contrasts with the increased effect found in our analysis (`r vasdem_text`). An additional analysis found that lipophilic statins were more harmful than hydrophilic statins in vascular dementia, potentially due to their ability to cross the blood brain barrier.

@@

Apart from statins, few studies examining a lipid-regulating agent have been reported. One of the few classes for which data was available were fibrates, for minimal evidence of an effect on all-cause dementia was identified, [@ancelin2012] inconsistent with our finding of a small increase in all-cause dementia risk in those prescribed a fibrate.

To our knowledge, there is no previous study of the effect of preventative treatment with ezetimibe on any dementia outcome, and so we cannot compare our unexpected finding for that treatment.

@@

However, the findings of our analysis are subject to several limitations. There is a strong possibility of differential misclassification of dementia-related condition based on the exposure, as those with memory complaints are more likely to be classified as vascular dementia than Alzheimer's disease if their medical records contained prescriptions for lipid-regulating agents. Further, there is a potential for non-differential misclassification of the outcome based on the use of electronic health records to identify dementia cases.[@wilkinson2018; @mcguinness2019validity]

Our study is likely to be subject to confounding by indication, which could provide a potential explanation for the observed increased risk of vascular and other dementia with lipid regulating agent use. Supporting evidence for this interpretation comes from a variety of sources, including the results of the control outcome analyses. The slight harmful effect for the backpain outcome is substantially smaller than that observed for the ischemic heart disease outcome, indicating that the majority of the uncontrolled confounding is likely related to vascular factors. Additionally, we obtained the expected harmful result for Type 2 diabetes, where statins' mechanism of action on this outcome is not know to be vascular.[@collins2016a; @macedo2014] Further supporting evidence comes from the increasingly harmful effect when moving from the Probable/Possible AD to other dementia to vascular dementia outcomes, indicating that confounding by indication likely increases as the proportion of vascular cases in the outcome definition increases. A review of other available literature suggests that this observation (a harmful effect of lipid regulating agents on vascular-related outcome due to confounding by indication) is not unusual. Using a conventional epidemiological technique, a previous analysis also found an increased risk of coronary heart disease (analogous to the ischemic heart disease outcome used in our analysis) in those taking statins (`r estimate(1.31,1.04, 1.66, type = "HR")`).[@danaei2013b]

@@

We have provided new evidence on the potential repurposing of lipid-regulating agents for the prevention of all-cause dementia, Alzheimer’s disease, vascular dementia, and other dementia. We found limited evidence to support the use of lipid-regulating agents for the prevention of  probable or possible Alzheimer’s disease, but they were associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations were driven by those observed in the statin subgroup, which comprised the majority of participants in our cohort.

@@

>>>>

__Methods:__ A cohort study design using data from the CPRD, routinely collected between January 1995 and March 2016, was performed. Cox proportional hazard models, allowing for a time-varying treatment indicator, were used to estimate the association between seven lipid-regulating drug classes (vs. no drug) and five dementia outcomes (all-cause, vascular and other dementia, and probable and possible Alzheimer's disease).

__Results:__ We analyzed `r totaltext` with a total follow-up of `r comma(total_followup)` participant-years (median: `r fu_text`). We found little evidence that lipid-regulating agents were associated with risk of Alzheimer's disease (probable `r probad_text`; possible `r possad_text`), but there was evidence of an increased risk of all-cause (`r anydem_text`), vascular (`r vasdem_text`) and other dementia (`r othdem_text`). Evidence from a number of control outcomes (ischaemic heart disease `r ihd_text`; backpain `r backpain_text` and diabetes `r dm_type2_text`) indicated the presence of substantial residual confounding related to vascular factors.

@@

* Little evidence that lipid-regulating agents were associated with Alzheimer’s disease, but there was some evidence for a harmful effect on all-cause, vascular and other dementia. In all cases, the estimated associations were driven by the any statin subgroup, which comprised most participants in our cohort.

@@

Dementia is a major progressive neurocognitive disorder, the most common types of which are Alzheimer’s disease, vascular dementia and Lewy Body dementia.[@prince2016] Despite an increasing number of cases globally and decades of research, there remains much unknown about the pathogenesis and progression of the disease, and, at present, no effective treatment exists to arrest, slow or reverse the cognitive decline associated with the condition.[@jobke2018] Drug repurposing, the identification of new applications for previously approved drugs, may provide an efficient mechanism to discover new effective preventative and therapeutic treatments for dementia.[@ashburn2004; @pushpakom2019]

@@

We therefore aimed to test the association between several major classes of LRA and all-cause dementia, Alzheimer's disease, vascular dementia and other dementia, in the Clinical Practice Research Datalink (CPRD), a large, population-based electronic health record (EHR) database.[@herrett2015]

@@

We performed a cohort study using data from the CPRD. Our initial sample included all participants registered at a participating practice between 1 January 1995 and 29 February 2016 who had a flag for "research quality" data. Records pre-dating the 1995 cut-off were excluded from the analysis as data quality and reliability is thought to be higher after this date.[@wolf2019] All events of interest were identified using predetermined code lists, which are available for inspection (see [Data/code availability](#data-code-avail)).

@@

Participants were included in our study cohort if their record contained any of the following index events: a Read code for a diagnosis of hypercholesterolemia or related condition; a Read code for prescription of a LRA (such as statins); a total cholesterol test result of >4mmol/L; or a low density lipoprotein cholesterol (LDL-c) test result of >2mmol/L.

These index events allowed us to define a population of participants who were either at risk of hypercholesterolemia, as indicated by the elevated total or LDL-c test results, or had already been diagnosed with it, as indicated by a diagnostic code or related prescription. This approach, conditioning entry into the study on being either “at-risk” or already diagnosed with hypercholesterolemia, was employed to reduce confounding by indication that we would expect to observe in a general population cohort.

The index date for a participant was defined as the date where the first relevant code or test result (as detailed above) was recorded on their clinical record. Participants were followed up until the earliest of: an outcome of interest; death; end of follow-up (29 February 2016); last registration date with their GP practice; last CPRD collection date for their practice. Participants were removed from our sample if they were less than 40 years of age, had less than 12 months of "research quality" data prior to their index date, were simultaneously prescribed more than one LRA (due to the difficulty of assigning these patients to a single exposure group), or were diagnosed with dementia before or on the date of the index event.

@@

To address the potential for immortal time bias, we employed a time-varying indicator of treatment status to correctly allocate time-at-risk to the exposed and unexposed groups.[@levesque2010] Under this model, all participants entered the unexposed group on their index date and were moved into the exposed group on the date they were first prescribed an eligible LRA. Participants whose index event was a LRA prescription entered the study and the exposed group on the same day, and so contributed no time-at-risk to the unexposed group.

A participant’s drug class was assigned based on their first recorded prescription, and any drug switching was ignored to mimic an intention-to-treat approach. We did however tabulate how often the initial drug class was stopped (defined as last prescription of the primary class being followed by at least six months of observation), added to (defined as a second drug class being prescribed before the last prescription of the initial class), or switched (defined as a second drug class being prescribed after the last prescription of the initial class).

@@

All analyses were performed in STATA 15. Cox proportional hazard models with a time-varying treatment indicator were used to estimate the hazard ratio and corresponding 95% confidence intervals, allowing for potential clustering of outcomes by practice. Participant's age was used as the time axis for all models.[@lamarca1998; @gail2009; @pencina2007] To observe the effect of adjusting for additional covariates, we compared models adjusted for age only and age and sex to the fully adjusted model. Additional analyses stratified by outcome and drug class were also performed.

@@

We performed several sensitivity analyses. As statins are contraindicated in pregnancy,[@nationalinstituteforhealthandcareexcellence]  we ran the models described above but excluding participants below the age of 55. Given the different ability of lipophilic statins to cross the blood-brain barrier,[@sierra2011] we further stratified the statin exposure group into lipophilic (Atorvastatin, Lovastatin, Simvastatin, Cerivastatin) and hydrophilic (Pravastatin, Rosuvastatin, Fluvastatin) statins. Finally, we included three control outcomes with known associations with statin use.[@collins2016a; @herrett2021] Using the fully adjusted model, we investigated the association between LRA and back pain (negative control), ischaemic heart disease (positive protective control), and Type 2 diabetes (positive harmful control).

@@

A total of `r format(attrition[14,1], big.mark = ",")` participants met the inclusion criteria for our cohort (See Supplementary Figure 2 for the attrition flowchart), with a total follow-up of `r comma(total_followup)` patient years at risk. Most participants were included in the cohort due to an elevated test result (`r index_event_text`). The median age at index was `r age_text` and participants were followed up for a median of `r fu_text`. During follow-up, an all-cause dementia diagnosis was recorded for `r format(p1[1,2], big.mark = ",")` patients  (`r format(p1[2,2], big.mark = ",")` probable Alzheimer's disease, `r format(p1[3,2], big.mark = ",")` possible Alzheimer's disease, `r format(p1[4,2], big.mark = ",")` vascular dementia, `r format(p1[5,2], big.mark = ",")` other dementia). The distribution of baseline characteristics across the drug classes can be seen in Table \@ref(tab:cprdCharacteristics-table).

Most participants (`r percentage.statins`) prescribed a lipid-regulating agent were prescribed a statin. We excluded the "Ezetimibe and statins" and "Nicotinic acid groups" classes from subsequent analysis based on the extremely small number of participants in these groups (Table \@ref(tab:cprdCharacteristics-table)). The stopping, addition and switching of drug classes was common across all exposure groups (Supplementary Table 3).

@@

As shown in Figure \@ref(fig:cprdPrimary), our results show little evidence of an effect of lipid-regulating agents on probable (`r probad_text`) and possible (`r possad_text`) Alzheimer's disease when compared with no treatment, except for fibrates on probable Alzheimer's disease (`r probad_fib_text`).

@@

In contrast to the findings for Alzheimer's disease outcomes, lipid-regulating agents were associated with an increased risk of a subsequent diagnosis of vascular dementia (`r vasdem_text`) or other dementia (`r othdem_text`). Again, this effect was driven mainly by the any statin subgroup, but there was some evidence that ezetimibe was associated with an increased risk of vascular (`r vasdem_eze_text`) and other (`r othdem_eze_text`) dementia.

@@

Adjustment for additional covariates beyond age and sex had a limited impact (Supplementary Figure 3), except for the Probable Alzheimer's disease outcome, where the full adjustment attenuated to the null the protective effect observed when adjusting only for age and sex.

@@

For our control outcomes (Supplementary Figure 6), there was some evidence that patients prescribed a lipid-regulating agent had increased risks of back pain (`r backpain_text`), ischaemic heart disease (`r ihd_text`) and Type 2 diabetes (`r dm_type2_text`).

@@

There was little evidence that lipid-regulating agents had any observed effect on probable and possible Alzheimer's when compared with no treatment, but some evidence they were associated with an increased risk of an all-cause dementia, vascular dementia and other dementia diagnosis. The effect observed in each case was driven by the any statin subgroup, which included a substantial majority of participants. For the other drug classes, there was limited evidence of an association with any outcome, with two exceptions. Ezetimibe was associated with increased risk of vascular and other dementia, while fibrates were associated with increased risk of all-cause dementia and probable Alzheimer's disease.

@@

These sources of evidence conflict with the findings of our analysis, where statin use was associated with an increased risk of all-cause dementia (`r anydem_text`). However, some of the included studies in the meta-analysis specifically exclude vascular dementia from the definition of all-cause dementia,[@chao2015] which may lead to an artificial protective effect of statins on all-cause dementia and limit the ability for comparison between studies.

@@

Our results are broadly in line with the findings of two distinct approaches examining the effect of statin treatment on subsequent Alzheimer's disease. No randomized trials of statins for the prevention of Alzheimer's disease have been reported, but a recent meta-analysis of 20 observational studies found statins were associated with a reduced risk of Alzheimer's disease (RR 0.69, 95% CI 0.60–0.80) with stronger evidence than observed in our analysis.[@poly2020] Additionally, a recent Mendelian randomization study examining the effect of genetic inhibition of HMGCR on Alzheimer’s disease (a genetic proxy for statin treatment) provided equivocal evidence (OR: 0.91, 95%CI: 0.63-1.31) but was not inconsistent with our results.[@williams2020]

Our additional analysis found no difference in effect between lipophilic and hydrophilic statins for the prevention of Alzheimer's disease, consistent with a recent meta-analysis of observational studies.[@chu2018]

@@

Far fewer studies have tested the association between lipid-regulating agents and vascular dementia or other dementia. A recent review found four observational studies examining the association of statins and vascular dementia found limited evidence for an effect (RR:0.93, 95% CI 0.74–1.16).[@poly2020] This contrasts with the increased effect found in our analysis (`r vasdem_text`). An additional analysis found that lipophilic statins were more harmful than hydrophilic statins in vascular dementia, potentially due to their ability to cross the blood brain barrier.

@@

Apart from statins, few studies examining a lipid-regulating agent have been reported. One of the few classes for which data was available were fibrates, minimal evidence of an effect on all-cause dementia was identified,[@ancelin2012] inconsistent with our finding of a small increase in all-cause dementia risk in those prescribed a fibrate.

A previous Mendelian randomization study provided an estimate of the effect of ezetimibe on Alzheimer's disease (OR: 1.17, 95%CI: 0.73-1.87),[@williams2020] consistent with the finding from our analyses. To our knowledge, there is no previous study of the effect of preventative treatment with ezetimibe on risk of vascular dementia, and so we cannot compare our unexpected finding for that comparison.

<!----------------------------------------------------------------------------->
&nbsp;

@@

However, the findings of our analysis are subject to several limitations. There was a  possibility of differential outcome misclassification based on the exposure, as we cannot exclude the possibility that for people with memory complaints, a diagnosis of vascular dementia might be made more frequently than Alzheimer’s disease if their medical records contain prescriptions for lipid-regulating agents.

Further, there is a potential for non-differential misclassification of the outcome based on the use of electronic health records to identify dementia cases.[@wilkinson2018; @mcguinness2019validity]

Our study is likely to be subject to confounding by indication, which could provide a potential explanation for the observed increased risk of vascular and other dementia with lipid regulating agent use. Supporting evidence for this interpretation comes from a variety of sources, including the results of the control outcome analyses. The slight harmful effect for the backpain outcome is substantially smaller than that observed for the ischaemic heart disease outcome, indicating that the majority of the uncontrolled confounding is likely related to vascular factors. Additionally, we obtained the expected harmful result for Type 2 diabetes, where statins' mechanism of action on this outcome is unlikely to be vascular.[@collins2016a; @macedo2014] Further supporting evidence comes from the increasingly harmful effect when moving from the Probable/Possible Alzheimer's disease to other dementia to vascular dementia outcomes, indicating that confounding by indication likely increases as the proportion of cases with a vascular component increases. A review of other available literature suggests that this observation (a harmful effect of lipid regulating agents on vascular-related outcome due to confounding by indication) is not unusual. Using a conventional epidemiological technique, a previous analysis also found an increased risk of coronary heart disease (analogous to the ischaemic heart disease outcome used in our analysis) in those taking statins (`r estimate(1.31,1.04, 1.66, type = "HR")`).[@danaei2013b]

@@

We have provided new evidence on the association of lipid-regulating agent prescription with all-cause dementia, Alzheimer’s disease, vascular dementia, and other dementia. We found limited evidence to support the use of lipid-regulating agents for the prevention of  probable or possible Alzheimer’s disease, but they were associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations were driven by those observed in the any statin subgroup, which comprised the majority of participants in our cohort.

@@

<!-- TODO 11,119 participants removed because their endate was . This explains the discrepancy between the total number in this figure and the table above. -->

### 2021-07-27 manuscripts/cprd/supp-materials.Rmd
End of day:

  * Use of a time-varying treatment indicator, to correctly classify time-at-risk.
  * Inclusion of additional covariates that are adjusted for in the main model.

@@

## Supplementary Table 1: STROBE Reporting Guidelines {#supp-tab-strobe}

@@

__To be copied in__

@@

&nbsp;

@@

# Supplementary figures {#supp-fig}
## Supplementary Figure 1 {#supp-fig-outcome-decision}

@@

## Supplementary Figure 3 {#supp-fig-unadjusted}
```{r supp-fig-comparison,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome using three models adjusted for age, age and sex, and all covariates respectively."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_unadjusted.jpeg"))
```

@@

## Supplementary Figure 4 {#supp-fig-pregnancy}
```{r supp-fig-pregnancy,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome, removing participants who were less than 55 years of age at index."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_pregnancy.jpeg"))
```

@@

## Supplementary Figure 5 {#supp-fig-sta-type}
```{r supp-fig-sta-type, echo = FALSE, fig.cap="Association of statins with a dementia or related outcome, stratified by statin lipophilic/hydrophilic properties."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_sta_type.jpeg"))
```

@@

## Supplementary Figure 6 {#supp-fig-backpain}
```{r supp-fig-backpain,echo = FALSE, fig.cap="Association of any lipid regulating agent with backpain, ischemic heart disease (IHD), and Type 2 diabetes."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_control_outcomes.jpeg"))
```

>>>>

-   Use of a time-varying treatment indicator, to correctly classify time-at-risk.
-   Inclusion of additional covariates that are adjusted for in the main model.

@@

## Supplementary Table 1: RECORD Reporting Guidelines {#supp-tab-strobe}

@@

**To be copied in**

@@

@@

# Supplementary figures {#supp-fig}

## Supplementary Figure 1 {#supp-fig-outcome-decision}

@@

```{r unadjustedComparison, include = FALSE}
```

```{r pregnancy, include = FALSE}
```

```{r statinType, include = FALSE}
```

```{r controlOutcomesText, include = FALSE}
```

```{r controlOutcomes, include = FALSE}
```

## Supplementary Figure 3 {#supp-fig-unadjusted}

```{r supp-fig-comparison,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome using three models adjusted for age, age and sex, and all covariates respectively."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_unadjusted.png"))
```

@@

## Supplementary Figure 4 {#supp-fig-pregnancy}

```{r supp-fig-pregnancy,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome, removing participants who were less than 55 years of age at index."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_pregnancy.png"))
```

@@

## Supplementary Figure 5 {#supp-fig-sta-type}

```{r supp-fig-sta-type, echo = FALSE, fig.cap="Association of statins with a dementia or related outcome, stratified by statin lipophilic/hydrophilic properties."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_sta_type.png"))
```

@@

## Supplementary Figure 6 {#supp-fig-backpain}

```{r supp-fig-backpain,echo = FALSE, fig.cap="Association of any lipid regulating agent with backpain, ischemic heart disease (IHD), and Type 2 diabetes."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_control_outcomes.png"))
```
